Supplementary Information: Improving Monocular Depth Estimation with Global Depth Histogram Matching

### 2019-10-14 paper/cvpr2020/sections/hardware.tex
Updated method figure and supplement.

%We determined camera intrinsics and extrinsics for the Kinect's RGB camera and
%the scanning system using the standard camera calibration toolbox in MATLAB.
%In general it is impossible to compensate for the offset between camera and
%SPAD, but we do add the z displacement from spad to kinect to partially
%compensate for it.
%% \textcolor{red}{Mark, please write a short paragraph on calibration details, including any warping of the SPAD histograms you did to compensate for the offset in camera and SPAD position.}
%
%Missing:
%%
%\begin{itemize}
%\item RGB resolution used for MDE
%\item do we also have Kinect depth maps for comparison? (yes) kinect resolution: RGB is 1920x1080 and depth camera is 512x424
%\end{itemize}

>>>>

We determined camera intrinsics and extrinsics for the Kinect's RGB camera and
the scanning system using the standard camera calibration toolbox in MATLAB.
Once the SPAD volume has been summed over the spatial dimensions, it is unclear
how to warp the resulting 1D transient to the RGB camera's perspective (in a
single-pixel flash LiDAR setup, the $512 \times 512$ ground truth data is not
available). However, we do shift the 1D transient by the $z$ displacement
between the SPAD and the RGB camera to partially account for this.

%% \textcolor{red}{Mark, please write a short paragraph on calibration details, including any warping of the SPAD histograms you did to compensate for the offset in camera and SPAD position.}
%
%Missing:
%%
%\begin{itemize}
%\item RGB resolution used for MDE
%\item do we also have Kinect depth maps for comparison? (yes) kinect resolution: RGB is 1920x1080 and depth camera is 512x424
%\end{itemize}

### 2019-10-14 paper/cvpr2020/sections/method.tex
Updated method figure.
Updated method figure and supplement.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{full_pipeline.jpeg}
  \caption{\textbf{Overview of the full pipeline} We use a CNN to get an initial
  per-pixel depth estimate. Then we perform exact histogram matching using
  intensity-weighted pixel values on the corrected SPAD data.}
\end{figure}

@@

\subsection{Ambient Rejection and Falloff Correction}
To remove the ambient and dark count photons from the histogram, we apply a
preprocessing step. Assuming that the first $N$ bins of the SPAD histogram
contain only ambient and dark count photons (this is equivalent to assuming the
minimum image depth is larger than $N \cdot c\Delta T/2$), we estimate 
and dark count rate $b$ as $\hat b = \frac{1}{N}\sum_{n=0}^N h[n]$ We then
note that that shining a laser through a 2D diffuser and
measuring a single pixel is very physically similar to recording a single 
non-line-of-sight measurement. We
compute the absolute first difference $d[n] = \abs{h[n] - h[n+1]}$ and find 
the set of edges. Intuitively, given a sufficiently large value of $b$, the Poisson sample
$h[n] \sim \mathcal{P}(b)$ should behave roughly like a Gaussian with variance $b$, and the difference
$h[n] - h[n+1]$ should behave like a Gaussian with variance $2b$. Therefore, we
can detect edges that occur between bins that contain actual signal photons and
bins that do not by computing the set 
\begin{equation}
  \mathcal{E} = \mset{n}{d[n] > \alpha\sqrt{2\hat b}}.
  \label{eq:edge_set}
\end{equation}
In this work, we use $\alpha = 2$.
We identify the closest and furthest edges $n_{first} = \min \mathcal{E}$ and
$n_{last} = \max \mathcal{E}$.
From \cite{Xin2019} Proposition 3, these edges
correspond to the discontinuties in the depth map 
at the closest and furthest image depths, and therefore all of the signal
photons must be in the bins between these two edges.
We refine these estimates by
computing $n'_{first} = \max\set{n_{first}}\cup\mset{n}{h[n] > \hat b, n \leq n_{first}}$ and
$n'_{last} = \min\set{n_{last}}\cup\mset{n}{h[n] > \hat b, n \geq n_{last}}$
Finally, we reject all photons outside of the valid range by clamping those bins
to 0, and we subtract our ambient estimate from the photons within the valid
range, giving an ambient-free transient
\begin{equation}
  h'_{clean}[n] = \begin{cases}
    0 & n < n'_{first} \\
    \max(h[n] - \hat b, 0) & n'_{first} \leq n \leq n'_{last} \\
    0 & n > n'_{last} \\
  \end{cases}.
  \label{eq:h_clean}
\end{equation}
To compensate for the radiometric falloff, we scale each bin by its squared depth,
which is consistent with our hardware prototype's radiometric falloff model.
This gives
\begin{equation}
  h_{clean}[n] = h'_{clean}[n] \cdot r_n^2
  \label{eq:h_scaled}
\end{equation}
where $r_n= \paren*{n + \frac12}\paren*{\frac{c\Delta t}{2}}$ is the depth
corresponding to bin $n$
As a final preprocessing step, we re-bin the transient, applying the Spacing-Increasing Discretization
(SID) of \cite{Fu2018} to produce $h_{target}$. This reduces the number of bins in the histogram
matching procedure, giving speedups while maintaining performance. In this work,
we use 140 bins for the simulated results and 600 bins for the captured results.

@@

\begin{algorithm}[H]
 \caption{Pixel Movement} 
 \label{alg:ehm}
 \begin{algorithmic}
  \Procedure{FindMovement}{$h_s$ of length $M$, $h_t$ of length $N$} 
    \State Initialize $T$ as an $M \times N$ array.
    \For{$m$ in $1,\ldots,M$}
      \For{$n$ in $1,\ldots,N$}
        \State $p_s \gets \sum_{\ell=1}^{n-1} T[m, \ell]$
        \State $p_t \gets \sum_{\ell=1}^{m-1} T[\ell, n]$
        \State $T[m, n] \gets \min(h_s[m] - p_s, h_t[n] - p_t)$
      \EndFor
    \EndFor
    \State \textbf{return} $T$
  \EndProcedure
 \end{algorithmic}
\end{algorithm}

Finally, for a pixel $z_0[i, j]$ in depth bin $k$, we sample the new depth bin from the
distribution $T[k, :]/\sum_{n=1}^NT[k,n]$ (which is a distribution over target
depth bins $1,\ldots,N$) to get the bin of $\hat z[i,j]$. We can then apply the
standard SID decoding procedure to get the depth corresponding to that bin
\textcolor{red}{Does the above seem reasonable to include? Not sure if the
  pseudocode is helpful/necessary or not.}
%\subsection{Implementation Details}
%For the Monocular Depth Estimator, we use pretrained versions of the
%the Deep Ordinal Regression Network (DORN) \cite{} and the DenseDepth Network.
%The exact histogram matching method is as described in \cite{}.

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{method.pdf}
  \caption{\textbf{Overview of the full pipeline} We use a CNN to get an initial
  per-pixel depth estimate. Then we perform exact histogram matching using
  intensity-weighted pixel values on the corrected SPAD data.}
\end{figure}

@@

\subsection{Ambient Rejection and Falloff Correction}
To remove the ambient and dark count photons from the histogram, we apply a
preprocessing step. Assuming that the first $N$ bins of the SPAD histogram
contain only ambient and dark count photons (this is equivalent to assuming the
minimum image depth is larger than $Nc\Delta T/2$), we estimate 
and dark count rate $b$ as $\hat b = \frac{1}{N}\sum_{n=0}^N h[n]$ We then
note that that shining a laser through a 2D diffuser and
measuring a single pixel is very physically similar to recording a single 
non-line-of-sight measurement. We
compute the absolute first difference $d[n] = \abs{h[n] - h[n+1]}$ and find 
the set of edges. Intuitively, given a sufficiently large value of $b$, the Poisson sample
$h[n] \sim \mathcal{P}(b)$ should behave roughly like a Gaussian with variance $b$, and the difference
$h[n] - h[n+1]$ should behave like a Gaussian with variance $2b$. Therefore, we
can detect edges that occur between bins that contain actual signal photons and
bins that do not by computing the set 
\begin{equation}
  \mathcal{E} = \mset{n}{d[n] > \beta\sqrt{2\hat b}}.
  \label{eq:edge_set}
\end{equation}
In this work, we use $\beta = 5$.
We identify the closest and furthest edges $n_{first} = \min \mathcal{E}$ and
$n_{last} = \max \mathcal{E}$.
From \cite{Xin2019} Proposition 3, these edges
correspond to the discontinuties in the depth map 
at the closest and furthest image depths, and therefore all of the signal
photons must be in the bins between these two edges.
We refine these estimates by
computing $n'_{first} = \max\set{n_{first}}\cup\mset{n}{h[n] > \hat b, n \leq n_{first}}$ and
$n'_{last} = \min\set{n_{last}}\cup\mset{n}{h[n] > \hat b, n \geq n_{last}}$
Finally, we reject all photons outside of the valid range by clamping those bins
to 0, and we subtract our ambient estimate from the photons within the valid
range, giving an ambient-free transient
\begin{equation}
  h'_{clean}[n] = \begin{cases}
    0 & n < n'_{first} \\
    \max(h[n] - \hat b, 0) & n'_{first} \leq n \leq n'_{last} \\
    0 & n > n'_{last} \\
  \end{cases}.
  \label{eq:h_clean}
\end{equation}
To compensate for the radiometric falloff, we scale each bin by its squared depth,
which is consistent with our hardware prototype's radiometric falloff model.
This gives
\begin{equation}
  h_{clean}[n] = h'_{clean}[n] \cdot r_n^2
  \label{eq:h_scaled}
\end{equation}
where $r_n= \paren*{n + \frac12}\paren*{\frac{c\Delta t}{2}}$ is the depth
corresponding to bin $n$
As a final preprocessing step, we re-bin the transient, applying the Spacing-Increasing Discretization
(SID) of \cite{Fu2018} to produce $h_{target}$. This reduces the number of bins in the histogram
matching procedure, giving speedups while maintaining performance. In this work,
we use 140 bins for the simulated results and 600 bins for the captured results.

@@

\begin{algorithm}[H]
 \caption{Pixel Movement} 
 \label{alg:ehm}
 \begin{algorithmic}
  \Procedure{FindMovement}{$h_s$ of length $M$, $h_t$ of length $N$} 
    \State Initialize $T$ as an $M \times N$ array of zeros.
    \For{$m$ in $1,\ldots,M$}
      \For{$n$ in $1,\ldots,N$}
        \State $p_s \gets \sum_{\ell=1}^{n-1} T[m, \ell]$
        \State $p_t \gets \sum_{\ell=1}^{m-1} T[\ell, n]$
        \State $T[m, n] \gets \min(h_s[m] - p_s, h_t[n] - p_t)$
      \EndFor
    \EndFor
    \State \textbf{return} $T$
  \EndProcedure
 \end{algorithmic}
\end{algorithm}

Finally, once we have obtained $T$, for a pixel $z_0[i, j]$ in depth bin $k$, we sample the new depth bin from the
distribution $T[k, :]/\sum_{n=1}^NT[k,n]$ (which is a distribution over target
depth bins $1,\ldots,N$) to get the bin of $\hat z[i,j]$. We can then apply the
standard SID decoding procedure to get the depth corresponding to that bin
\textcolor{red}{Does the above seem reasonable to include? Not sure if the
  pseudocode is helpful/necessary or not.}
%\subsection{Implementation Details}
%For the Monocular Depth Estimator, we use pretrained versions of the
%the Deep Ordinal Regression Network (DORN) \cite{} and the DenseDepth Network.
%The exact histogram matching method is as described in \cite{}.

### 2019-10-14 paper/cvpr2020/sections/simulation.tex
Removed whitespace from comparison figure.
Updated method figure and supplement.

\begin{figure*}[!t]
  \includegraphics[width=\textwidth]{comparison/densedepth_468_comparison.png}
  \includegraphics[width=\textwidth]{comparison/densedepth_194_comparison.png}
  %\includegraphics[width=\textwidth]{comparison/densedepth_258_comparison.png}
  \caption{Simulated results from NYU v2 computed with the DenseDepth CNN. The depth maps estimated by the CNN are reasonable, but contain systematic error. Oracle access to the ground truth depth maps, either through the median depth or the depth histogram, can remove this error and correct the depth maps. The proposed method uses a single diffused SPAD and does not rely on ground truth depth, but it achieves a quality that closely matches the best-performing oracle.}
	\label{fig:results_simulated}
\end{figure*}

@@

We use the NYU Depth v2 dataset to evaluate our method. This dataset consists of
249~training and 215~testing scenes of RGB-D data captured using a Microsoft
Kinect.
We compute the standard intensity image (i.e. the Y
component of the YCbCr image, scaled to be in $[0,1]$) from the RGB input. We compute the weighted depth
histogram of the ground truth depth image using these intensity weights, and sum
over the spatial dimension to get a 1D histogram $h$. We multiply each bin by
$1/d^2$, which accounts for the radiometric falloff, and convolve with the SPAD
jitter kernel, which we model as a standard Gaussian with a full-width
half-maximum of 70ps. We scale this histogram by a signal photon count (set at
$s = 10^6$) and then add a fixed number of photons $b \in \set{10^5, 2\times 10^4, 10^4}$ evenly distributed across all
bins to simulate the ambient and dark count photons, modeling
signal-to-background ratios (SBR) of $10, 50$ and $100$ respectively. Finally,
we Poisson sample each bin to get the final simulated SPAD transient.
% \textcolor{red}{Need more details: how did you simulate the SPAD measurements from this: discuss albedo, square distance falloff, Poisson noise, background, number of SPAD bins, define what SBR is etc.}

>>>>

\begin{figure*}[!t]
  % \includegraphics[width=\textwidth]{comparison/densedepth_468_comparison.pdf}
  % \includegraphics[width=\textwidth]{comparison/densedepth_194_comparison.pdf}
  \includegraphics[width=\textwidth]{comparison.pdf}
  %\includegraphics[width=\textwidth]{comparison/densedepth_258_comparison.png}
  \caption{Simulated results from NYU v2 computed with the DenseDepth CNN. The depth maps estimated by the CNN are reasonable, but contain systematic error. Oracle access to the ground truth depth maps, either through the median depth or the depth histogram, can remove this error and correct the depth maps. The proposed method uses a single diffused SPAD and does not rely on ground truth depth, but it achieves a quality that closely matches the best-performing oracle.}
	\label{fig:results_simulated}
\end{figure*}

@@

We use the NYU Depth v2 dataset to evaluate our method. This dataset consists of
249~training and 215~testing scenes of RGB-D data captured using a Microsoft
Kinect.
We compute the standard intensity image (i.e. the Y
component of the YCbCr image, scaled to be in $[0,1]$) from the RGB input. We compute the weighted depth
histogram of the ground truth depth image using these intensity weights, and sum
over the spatial dimension to get a 1D histogram $h$. We multiply each bin by
$1/z^2$, which accounts for the radiometric falloff, and convolve with the SPAD
jitter kernel, which we model as a standard Gaussian with a full-width
half-maximum of 70ps. We scale this histogram by a signal photon count (set at
$s = 10^6$) and then add a fixed number of photons $b \in \set{10^5, 2\times 10^4, 10^4}$ evenly distributed across all
bins to simulate the ambient and dark count photons, modeling
signal-to-background ratios (SBR) of $10, 50$ and $100$ respectively. Finally,
we Poisson sample each bin to get the final simulated SPAD transient (neglecting
pile-up).
% \textcolor{red}{Need more details: how did you simulate the SPAD measurements from this: discuss albedo, square distance falloff, Poisson noise, background, number of SPAD bins, define what SBR is etc.}
